The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2020-05-07 Paper/jaccardbaetica.tex
review xavi 2draft

%\xavi{queda raro esto de "en arqueología la cultura material nos permite ver la distribución en el Mediterráneo". Yo iría directa a hablar del Mediterráneo sin la primera frase.}

@@

%\xavi{este aim es demasiado general}
%This paper aims to understand the production dynamics in connection with different areas within the Roman Empire.

\xavi{le das demasiada importancia al similarity index. Es una simple herramienta de tu trabajo, y no un objetivo}.
This paper aims to study the olive oil market connection between province by calculating the similarity index. Specifically, our work pretends to detect microeconomic processes focused on a commercial product from a specific province \citep{isaksen_network_2006}. We want to understand the pattern of olive oil production linked to amphora workshops and amphoric stamps used to mark them. We focus here on exploring the economic relation between stamps and amphora production and distribution centres.

Two case studies have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}). In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop; in Roman provinces, our aim is detecting groups of stamps concentrated in an area or if some groups have an important role for the exportation of olive oil in those provinces. This economic connection could be identified by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamps in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific Roman province. 
%\xavi{affected? No sería "identified"? Es decir, la correlación es la evidencia, pero la causa sería la interacción económica...no?}

%Our case study examines the relationship between the distribution of amphoric stamps and the workshops.
In particular, we study the distribution of amphoric stamp in order to identify a correlation between geographical distance and amphoric stamps. Based on this assumption, we proposed three hypotheses: a) we can identify a correlation between spatial distance and the distribution of stamps, b) stamps located in close workshops share similar traits and c) Low mobility of amphoric stamps to other regions: stamps always stay in the same region.

@@

%\xavi{esta frase tendría que ir al final de la Intro; si pones subsecciones entonces queda muy raro creo}

@@

%\xavi{creo que le falta chicha aquí, sobre qué se ha hecho hasta ahora, qué quieres saber específicamente, el caso de estudio...}

%\xavi{Yo aplicaría aquí el feedback de los mapas para lo del libro...no creo que colorearlo basándose en el DEM aporte demasiado no?}

\section{Britannia and Germania}
\xavi{este título de sección debería cambiarse por algo más descriptivo. Case studies o algo asínnn? Por otra parte, igual podrías dividirlo en 2 subsections}

%\xavi{Esta parte más histórica está chachi pero creo que debería estar en la Intro. Ahí podrías explicar la organización, y a partir de estas fuentes escritas plantear preguntas e hipótesis que justifiquen el paper y tu análisis usando arqueología.}

It is well known that this product was not frequently consumed by indigenous population \xavi{CITATION NEEDED}. The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of a high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[ 1]{carreras_britannia_1998}. Additionally, the land of \textit{Britannia} showed inadequate for the olive oil production due to low environmental conditions. This issue was necessary for creating an important export apparatus from \textit{Baetica} to supply Roman army.

At this moment, we detect an increase of the olive oil exportation concurring with the displacement of legions during the military campaigns \citep[161]{monfort_britanniaen_1998}. This fact will be spatial incidence in sites close at Hadrian Wall.  
Olive oil production in \textit{Baetica} would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic points \citep{carreras_atlantic_2012}. The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges. Thus, the network was mainly focused on the support of soldiers during the military campaigns. 
%debate intercambios puntuales o redes comerciales de abastecimiento

\section{A potential indicator of Roman economy: Dressel 20 stamps}
\xavi{no sería mejor que esta sección esté antes de hablar de los casos de estudios? Así, podrías ya hablar de sellos encontrados en germania y britannia}

@@

\section{Method and Analysis}
\xavi{estas 2 cosas nunca van juntas. En Method explicas las técnicas, y Analysis en general no se usa porque es el todo del trabajo}

The assumption of this study is analysing the effect of the production patterns between different centres \xavi{esto no es ninguna asunción no??}. We are specially interested in identifying links between producers and consumption centres using amphoric stamps. To do that, we use the CEIPAC database to collect the stamps from different places. The CEIPAC dataset contains over 50.000 of epigraphy records found in amphorae, mostly from \textit{Monte Testaccio}. Therefore, this study proposes a robust baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps. A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps. Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance. 
%\xavi{no veo la parte evolutiva en el paper; por otra parte, si es relevante no la lanzaría aquí al vuelo; hace falta explicar por qué es necesaria, qué implica, etc.}

@@

%\xavi{a ver, esto queda raro...si esto es una exploración de la base de datos entonces tiene que ir dentro de "Materials", así mientras vas comentando los sesgos y cosas generales de los datos (como soporte, pero no hace falta subsubsección "EDA")}

We calculated the frequency of distribution in amphora stamps with the aim to carry out an exploratory analysis (EDA) to know the distribution and the number of stamps per each centre.

@@

%The type of distribution is also frequent in the Roman economy where we observed a self-organised complex system pattern typical of a free market 
%\xavi{quieres decir que esto es una power base? yo creo que no, no es tan bestia no? Yo lo matizaría}

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/frequencystamp.pdf}
\caption{Histogram with the number of stamps (X axis) and the frequency of distribution per workshop (Y axis). The distribution is widely diverse with most workshops concentrated one stamp}
\label{stamps}
\end{figure}

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/frequency}
\caption{Distribution of the number of different code stamps (X axis) for each \textit{conventus} (Y axis). Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue)}
\label{frequency}
\end{figure}

\xavi{vale pero qué es cada punto? un workshop?}

@@

\subsection{Measuring the dissimilarity in stamps}

\xavi{en stamps o en yacimientos? o en provincias?}

@@

\subsection{Reception centres: Britannia and Germania}
\xavi{mejor consumption?}

@@

Most military sites that showed greater similarity in the results were geographically close. We identified two factors in the results: 1) nearby sites tend to share similar stamps and 2) most sites have different stamps. 
By contrast, most sites did not show a strong similarity in stamp correlation. Neither we found a grouping of similar stamps in a specific place. In general, we did not observe any production pattern that indicates a clear organisation between production centres and reception centres. In other words, our results did not indicate the presence of a specific production center from \textit{Baetica} in \textit{Britannia} province. Rather, it seems that olive oil production was distributed by non-specific production centres.

>>>>

@@

%\xavi{le das demasiada importancia al similarity index. Es una simple herramienta de tu trabajo, y no un objetivo}.
This paper aims to study the olive oil market connection between provinces by calculating the similarity of stamps. Specifically, our work pretends to detect microeconomic processes focused on a commercial product from a specific province \citep{isaksen_network_2006}. We want to understand the pattern of olive oil production linked to amphora workshops and amphoric stamps used to mark them. We focus here on exploring the economic relation between stamps and amphora production and distribution centres.

Two case studies have been studied in order to analyse the relation between production centres (\textit{Baetica}) and consumption centres (\textit{Britannia} and \textit{Germania}). In the case of \textit{Baetica} province, we want to identify the role of the stamps in the organisation of the workshop; in Roman provinces, our aim is detecting groups of stamps concentrated in an area or if some groups have an important role for the exportation of olive oil in those provinces. This economic connection could be identified by different aspects: a) correlation between spatial distance and centres based on the idea that closer workshops concentrate similar amphoric stamps in a specific area than the distant workshops and b) groups of similar stamps were concentrated in a specific Roman province.

In particular, we study the distribution of amphoric stamp in order to identify a correlation between geographical distance and amphoric stamps. Based on this assumption, we proposed three hypotheses: a) we can identify a correlation between spatial distance and the distribution of stamps, b) stamps located in close workshops share similar traits and c) Low mobility of amphoric stamps to other regions: stamps always stay in the same region.

@@

@@

\section{A potential indicator of Roman economy: Dressel 20 stamps}
%\xavi{no sería mejor que esta sección esté antes de hablar de los casos de estudios? Así, podrías ya hablar de sellos encontrados en germania y britannia}

@@

\section{Case study: consumption centers}
%\xavi{este título de sección debería cambiarse por algo más descriptivo. Case studies o algo asínnn? Por otra parte, igual podrías dividirlo en 2 subsections}

\subsection{Britannia}

It is well known that this product was not frequently consumed by indigenous population \citep[161]{monfort_britanniaen_1998}. The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of a high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[ 1]{carreras_britannia_1998}. Additionally, the land of \textit{Britannia} showed inadequate for the olive oil production due to low environmental conditions. This issue was necessary for creating an important export apparatus from \textit{Baetica} to supply Roman army.

At this moment, we detect an increase of the olive oil exportation concurring with the displacement of legions during the military campaigns \citep[161]{monfort_britanniaen_1998}. This fact will be spatial incidence in sites close at Hadrian Wall.  
Olive oil production in \textit{Baetica} would cross the Atlantic until they reached the province and redistribute throughout the area from a series of strategic points \citep{carreras_atlantic_2012}. The increase of the exportation of Dressel 20 amphorae created an important commercial network for exchanges. Thus, the network was mainly focused on the support of soldiers during the military campaigns.

\subsection{Germania}

\section{Material}

%\xavi{estas 2 cosas nunca van juntas. En Method explicas las técnicas, y Analysis en general no se usa porque es el todo del trabajo}

The goal of this study is analysing the effect of the production patterns between different centres. We are specially interested in identifying links between producers and consumption centres using amphoric stamps. To do that, we use the CEIPAC database to collect the stamps from different places. The CEIPAC dataset contains over 50.000 of epigraphy records found in amphorae, mostly from \textit{Monte Testaccio}. Therefore, this study proposes a robust baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps. A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps. Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.

@@

We calculated the frequency of distribution in amphora stamps with the purpose to carry out an exploratory analysis (EDA) to know the distribution and the number of stamps per each centre.

@@

%The type of distribution is also frequent in the Roman economy where we observed a self-organised complex system pattern typical of a free market

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/frequencystamp.pdf}
\caption{Histogram with the number of stamps (X axis) and the frequency of workshop (Y axis). The distribution is widely diverse with most workshops have only one stamp}
\label{stamps}
\end{figure}

@@

\begin{figure}[htp]
	\centering
\includegraphics[width=\linewidth]{figs/frequency}
\caption{Distribution of the number of different code stamps (X axis) for each \textit{conventus} (Y axis). Each dot corresponds to a workshop sorted by different areas. Colors are represented by areas divided into Hispalis (red), Astigi (green) and Corduba (blue)}
\label{frequency}
\end{figure}

%\xavi{vale pero qué es cada punto? un workshop?}

@@

\section{Method: Measuring the dissimilarity}

%\xavi{en stamps o en yacimientos? o en provincias?}

@@

\subsection{Consumption centres: Britannia and Germania}
%\xavi{mejor consumption?}

@@

Most military sites that showed greater similarity in the results were geographically close. We identified two factors in the results: 1) nearby sites tend to share similar stamps and 2) most sites have different stamps. 
By contrast, most sites did not show a strong similarity in stamp correlation. Neither we found a grouping of similar stamps in a specific place. In general, we did not observe any production pattern that indicates a clear organisation between production centres and consumption centres. In other words, our results did not indicate the presence of a specific production center from \textit{Baetica} in \textit{Britannia} province. Rather, it seems that olive oil production was distributed by non-specific production centres.
